# System Setup

## Quick start

If you can't be bothered reading simply do this:

1. [click here](https://www.rstudio.com/products/rstudio/download/#download)
2. open the file and click install
3. open RStudio
4. copy & paste the following lines to the command line:

```{r, eval=FALSE, message=FALSE, warning=FALSE}
install.packages("devtools")
```

For a bit more info keep reading...

## Bio-Linux

Setting up your system with is often one of the most time consuming and frustrating parts of starting out in bioinformatics. To avoid the hassle we'll be running [Bio-Linux](http://environmentalomics.org/bio-linux/) from a USB stick. This provides you with a Linux workstation pre-setup with many common bioinformatic tools.

## R

In addition to command line programs we'll be using R to plot figures. R is a statistical programming language commonly used for down-stream statistical analysis and visualisation. R can be run from the command line, or using the popular user-interface [RStudio](https://www.rstudio.com/). 

### RStudio

We'll be using RStudio for the course, which isn't pre-installed on Bio-Linux - so simply [click here](https://www.rstudio.com/products/rstudio/download/#download) to download for the Bio-Linux system, open the file and install. 

### R packages

R packages 

[devtools](https://cran.r-project.org/web/packages/devtools/index.html)
```{r, eval=FALSE, message=FALSE, warning=FALSE}
install.packages("devtools")
```

[BiocInstaller](http://bioconductor.org/packages/BiocInstaller)
```{r, eval=FALSE, message=FALSE, warning=FALSE}
source('https://bioconductor.org/biocLite.R')
biocLite('BiocInstaller')
```

[DESeq2](https://bioconductor.org/packages/DESeq2)
```{r, eval=FALSE, message=FALSE, warning=FALSE}
source("https://bioconductor.org/biocLite.R")
biocLite("DESeq2")
```
